---
title: "SElshahawy_Assign8"
author: "Salma Elshahawy"
date: "3/17/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question_11, page 303, Introduction to probability

A company buys 100 lightbulbs, each of which has an **exponential lifetime of 1000 hours**. What is the expected time for the first of these bulbs to burn out?


```{r}
n <- 100
meu <- 1000
E_m <- meu/n # from excersie 10
E_m
```

## Question_14, page 303, intruduction to probability

Assume that $X_1$ and $X_2$ are independent random variables, each having an
exponential density with parameter $\lambda$. Show that $Z = X_1-X_2$ has density
$f_Z(z) = (1/2)e^{-\lambda|z|}$.


$f_Z(z) = (1/2)e^{-\lambda|z|}$ can be re-written as $f_Z(z) = \begin{cases} (1/2)e^{-\lambda z}, & \mbox{if } z \ge 0, \\ (1/2)e^{\lambda z}, & \mbox{if }z <0. \end{cases}$

Since $X_1$ and $X_2$ have exponential density, their PDF is

<CENTER>
$f_{X_1}(x)=f_{X_2}(x)=\begin{cases} \lambda e^{-\lambda x}, & \mbox{if } x\ge 0, \\ 0, & \mbox{otherwise. }\end{cases}$
</CENTER>

$$
\begin{split}
  f_Z(z) &= f_{X_1+(-X_2)}(z) \\
    &= \int_{-\infty}^{\infty} f_{-X_2}(z-x_1) f_{X_1}(x_1) dx_1 \\
    &= \int_{-\infty}^{\infty} f_{X_2}(x_1-z) f_{X_1}(x_1) dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda e^{-\lambda(x_1-z)} \lambda e^{-\lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{-\lambda x_1 + \lambda z} e^{-\lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{\lambda z - \lambda x_1 - \lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1
\end{split}
$$

Consider $z=x_1-x_2$, then $x_2=x_1-z$.

If $z \ge 0$, then $x_2=(x_1-z) \ge 0$, and $x_1 \ge z$, and, using WolframAlpha, $f_Z(z) = \int_{z}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1 = \frac{1}{2} \lambda e^{-\lambda z}$. 

If $z < 0$, then $x_2=(x_1-z) \ge 0$, and $x_1 \ge 0$, and $f_Z(z) = \int_{0}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1 =\frac{1}{2} \lambda e^{\lambda z}$.

Combining two sides we get $f_Z(z) = \begin{cases} (1/2)e^{-\lambda z}, & \mbox{if } z \ge 0, \\ (1/2)e^{\lambda z}, & \mbox{if }z <0. \end{cases}$

## Question_1, page 320,321

Let $X$ be a continuous random variable with mean $\mu = 10$ and variance
$\sigma^2 = 100/3$. Using Chebyshev's Inequality, find an upper bound for the
following probabilities.

a. $P(|X - 10| \ge 2)$
b. $P(|X - 10| \ge 5)$
c. $P(|X - 10| \ge 9)$
d. $P(|X - 10| \ge 20)$

Chebyshev Inequality: $P(|X-\mu|\ge\epsilon) \le \frac{\sigma^2}{\epsilon^2}$ or, per example 8.4, $P(|X-\mu|\ge k\sigma) \le \frac{1}{k^2}$.

Per problem, $\mu=10$ and $\sigma = \sqrt{\frac{100}{3}} = \frac{10}{\sqrt{3}}$.

If $\epsilon = k\sigma$, then $k=\frac{\epsilon}{\sigma} = \frac{\epsilon\sqrt{3}}{10}$.

Let $u$ be upper bound in Chebyshev's Inequality, then $u = \frac{1}{k^2} = \frac{1}{(\epsilon\sqrt{3}/10)^2} = \frac{100}{3\epsilon^2}$.

a. $\epsilon=2$, the upper bound is $u= \frac{100}{3\times 2^2} = 8.3333$. Since probability cannot be greater than $1$, the upper bound is $1$.
b. $\epsilon=5$, the upper bound is $u= \frac{100}{3\times 5^2} = 1.3333$. Since probability cannot be greater than $1$, the upper bound is $1$.
c. $\epsilon=9$, the upper bound is $u= \frac{100}{3\times 9^2} = 0.4115$.
d. $\epsilon=20$, the upper bound is $u= \frac{100}{3\times 20^2} = 0.0833$.


