---
title: "SElshahawy_Assign12"
author: "Salma Elshahawy"
date: "4/23/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset importing

```{r}
dataset = read.csv('who.csv', header = TRUE)
head(dataset)
```

The dataset is 190 instances(observations) and 10 attributes(features) as the following:

Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate TotExp: sum of personal and government expenditures.

```{r}
str(dataset)
summary(dataset)
```

## 1. 

Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}
library(ggplot2)
ggplot(dataset, aes(x = TotExp, y  = LifeExp)) + 
  geom_point()
```


Simple linear regression

```{r}
regressor = lm(LifeExp ~ TotExp, data = dataset)
summary(regressor)
```

P-value: is less than 0.05 significance level (2e-16) and F-Statistic is 65.26, so we can be sure that TotExp independent variable affecting the LifeExp. We can reject the null hypothesis. However, the R-squared is low (25.3%), which means that the model would explain only 25% of the model variability.

Std. Error is an estimate of the standard deviation of the coefficient, the amount it varies across cases. 

This model can not be presented by a simple linear regression. The Adjusted R-squared percent is too low, may be if we included more independent variables we can get a higher value. 


## 2. 

Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

```{r}
# raise LifeExp to 4.6 power
dataset$LifeExp_4.6 <- (dataset$LifeExp)^4.6
# raise TotExp to 0.06 power 
dataset$TotExp_0.06 <- (dataset$TotExp)^0.06
head(dataset)
```


```{r}
ggplot(dataset, aes(x = TotExp_0.06, y  = LifeExp_4.6)) + 
  geom_point() 
```


```{r}
regressor_2 = lm(LifeExp_4.6 ~ TotExp_0.06, data = dataset)
summary(regressor_2)
```

```{r}
ggplot(data = dataset, aes(x = TotExp_0.06, y = LifeExp_4.6)) +
geom_point() +
stat_smooth(method = "lm", col = "dodgerblue3") +
theme(panel.background = element_rect(fill = "white"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Linear Model Fitted to Data")
```

```{r}
plot(regressor_2$fitted.values, regressor_2$residuals,
     xlab = 'fitted values',
     ylab = 'residuals',
     main = 'Residual plot')
abline(h=0)
```

```{r}
# Residuals Q-Q plot
qqnorm(regressor_2$residuals)
qqline(regressor_2$residuals)
```

P-value: is still less than 0.05 significance level (2e-16) ,but F-Statistic got higher 507.7, so we can be sure that TotExp independent variable affecting the LifeExp. We can reject the null hypothesis. Moreover, the R-squared got better as well (72.8%), which means that the model would explain about 72.8% of the model variability.

Std. Error is an estimate of the standard deviation of the coefficient, the amount it varies across cases. 


This model Definitly is better than the previous one.  


## 3. 

Using the results from 2, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.

```{r}
predict(regressor_2, data.frame(TotExp_0.06=c(1.5, 2.5)), interval = 'predict')^(1/4.6)
```


## 4. 

Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

```{r}
regressor_3 <- lm(LifeExp ~ PropMD + TotExp + TotExp:PropMD, data = dataset)
summary(regressor_3)
```

```{r}
plot(regressor_3$fitted.values, regressor_3$residuals,
     xlab = 'fitted values',
     ylab = 'residuals',
     main = 'Residual plot')
abline(h=0)
```


```{r}
# Residuals Q-Q plot
qqnorm(regressor_2$residuals)
qqline(regressor_2$residuals)
```


Residual standard error is 8.765 and F-statistic is 34.49. Considering that average life expectancy is 67.38, the SE is not terrible and F-statistics is fairly high (but lower than in the first model). R2 is only 0.3574, so the model explains only 35.74% of variability, which is not high. P-value is nearly 0, so the relationship is not due to random variation.

Looking at residuals plots it is clear that there is no constant variability and that residuals are not normally distributed. This is not a good model to describe the relationship. Kind of similar to the first model.


## 5. 

Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?


```{r}
predict(regressor_3, data.frame(PropMD = 0.03, TotExp = 14), interval = 'predict')
```

The prediction is 107.696 years with 95% confidence interval between 84.247 and 131.144. The prediction could be realistic but it is very rare. I don't think that it is a good model, The Adjusted R-squared in way too low to explain the variability of the model.  

































