---
title: "SElshahawy_Assign12"
author: "Salma Elshahawy"
date: "4/23/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset importing

```{r}
library(RCurl)
dataset = read.csv('who.csv', header = TRUE)
head(dataset)
```

The dataset is 190 instances(observations) and 10 attributes(features) as the following:

Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate TotExp: sum of personal and government expenditures.

```{r}
str(dataset)
summary(dataset)
```

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}
library(ggplot2)
ggplot(dataset, aes(x = TotExp, y  = LifeExp)) + 
  geom_point()
```


Simple linear regression

```{r}
regressor = lm(LifeExp ~ TotExp, data = dataset)
summary(regressor)
```

P-value: is less than 0.05 significance level, so we can be sure that TotExp independent variable affecting the LifeExp.

F-Statistic: should be used in combination with P-value to decide the significance level. In this example, we have only one independent variable - F-statistic value doesn't imply anything.

R-squared: is the fraction of the variation in the dependent variable that is accounted for the independent variables.

Std. Error is an estimate of the standard deviation of the coefficient, the amount it varies across cases.


This model can not be presented by a simple linear regression. The Adjusted R-squared percent is too low, may be if we included more independent variables we can get a higher value. 

As The below graph illusterated, the TotExp has a non-linear relationship with the LifeExp dependent variable.


```{r}
ggplot(dataset, aes(x = TotExp, y  = LifeExp)) + 
  geom_point() +
  geom_smooth(method = lm)
```





















