---
title: "Untitled"
author: "Salma Elshahawy"
date: "4/14/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Linear Regression

In this example, I would demonestrate how to create a simple linear regression. Furthermore, I would make a residual dignoses to weigh our model quality and how far does it fit to the problem. 

The dataset I have is basically consists of one independent variable(year of experience) and one dependant variable (Salary). I would build a linear regression model to predict the salary at certain point. The relation between two variables is linear. 

## Importing the dataset
```{r}
library(RCurl)

myfile = getURL('https://raw.githubusercontent.com/salma71/Machine-learning-tutorial/master/Part%202%20-%20Regression/Section%204%20-%20Simple%20Linear%20Regression/Salary_Data.csv', ssl.verifyhost=FALSE, ssl.verifypeer=FALSE)

dataset = read.csv(textConnection(myfile), header = TRUE)
head(dataset)
```

## Splitting the dataset into the Training set and Test set
```{r}
# install.packages('caTools')
library(caTools)
set.seed(123)
split = sample.split(dataset$Salary, SplitRatio = 2/3)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)
```


## Fitting Simple Linear Regression to the Training set
```{r}
regressor = lm(formula = Salary ~ YearsExperience,
               data = training_set)
summary(regressor)
```

The model says that for a salary to increase one unit it required to increase years of experience variable by 9365 units. Also, we can see from the model, the Adjusted R-squared is 0.96. Means that the model can predict 96% of the data - which is good. 

## Residual dignosis

```{r}
par(mfrow=c(2,2)) # Split the plotting panel into a 2 x 2 grid
plot(regressor) # Plot the model information
```


The first plot (residuals vs. fitted values) is a simple scatterplot between residuals and predicted values. It look a bit random but so far equally distributed around the center.

The second plot (normal Q-Q) is a normal probability plot.  It gives so far a straight line which means that the errors are distributed normally.

The third plot (Scale-Location), like the the first, looks random.  No patterns. However, it gave me a strange inversed v-shape.

The last plot (Cook’s distance) tells us which points have the greatest influence on the regression (leverage points). We see that points 7, 9 and 12 have great influence on the model.
