---
title: "Inverse of a Matrix-Discussion week 2"
author: "Salma Elshahawy"
date: "2/4/2020"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

I tried to solve the problem C19 - page 205 - A First course in Linear Algebra by Rober A. Beezer. 

The question asks to find the inverse of a matrix if exits and verifying it.  

I solved the problem using hand and verify the solution using matlib package in r to get the inverse. 

We know that AI = inv(A)

## 1

We need to multiply by the Identy matrix to find the inverse in an augmented matrix

$$\left\{ { \begin{matrix} 1 & 3 & 1 \\ 0 & 2 & 1 \\ 2 & 2 & 1 \end{matrix} }|{ \begin{matrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{matrix} } \right\} \quad $$

## 2

Determine the pivot entry point, which is in this case is 1. We need to eleminate elements under the pivot using the Upper triangle and lower triangle method. 

$${ pivot }_{ 1 }\quad =\quad 1\\ { multiplier }_{ 1 }\quad =\quad \frac { 2 }{ 1 } \\ R3\quad =\quad R3\quad -\quad { multiplier }_{ 1 }\quad \times \quad R1\\ { pivot }_{ 2 }\quad =\quad 2\\ { multiplier }_{ 2 }\quad =\quad \frac { -4 }{ 2 } \\ R3\quad =\quad R3\quad -\quad { multiplier }_{ 1 }\quad \times \quad R1\\ \left\{ { \begin{matrix} 1 & 3 & 1 \\ 0 & 2 & 1 \\ 0 & -4 & -1 \end{matrix} }|{ \begin{matrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ -2 & 0 & 1 \end{matrix} } \right\} \quad \Longrightarrow \quad \left\{ { \begin{matrix} 1 & 3 & 1 \\ 0 & 2 & 1 \\ 0 & 0 & 1 \end{matrix} }|{ \begin{matrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ -2 & 2 & 1 \end{matrix} } \right\} $$

$${ pivot }_{ 3 }\quad =\quad 1$$

## 3

We eliminate the elements above the pivots  - Lower triangle. 

$$\left\{ { \begin{matrix} 1 & 3 & 1 \\ 0 & 2 & 1 \\ 0 & 0 & 1 \end{matrix} }|{ \begin{matrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ -2 & 2 & 1 \end{matrix} } \right\} \quad \Longrightarrow \quad \left\{ { \begin{matrix} 1 & 0 & \frac { -1 }{ 2 }  \\ 0 & 2 & 1 \\ 0 & 0 & 1 \end{matrix} }|{ \begin{matrix} 0 & \frac { -3 }{ 2 }  & 0 \\ 0 & 1 & 0 \\ -2 & 2 & 1 \end{matrix} } \right\} $$

$$divide\quad each\quad row\quad by\quad the\quad pivot\quad coeff.\\ \Longrightarrow \quad \left\{ { \begin{matrix} 1 & 0 & \frac { -1 }{ 2 }  \\ 0 & 2 & 0 \\ 0 & 0 & 1 \end{matrix} }|{ \begin{matrix} 1 & \frac { -3 }{ 2 }  & 0 \\ 2 & -1 & -1 \\ -2 & 2 & 1 \end{matrix} } \right\} \quad \Longrightarrow \quad \left\{ { \begin{matrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{matrix} }|{ \begin{matrix} 0 & \frac { -1 }{ 2 }  & \frac { 1 }{ 2 }  \\ 2 & -\frac { 1 }{ 2 }  & -\frac { 1 }{ 2 }  \\ -2 & 2 & 1 \end{matrix} } \right\} $$


$$The\quad inverse\quad should\quad be\quad \\ \\ \left[ \begin{matrix} 0 & \frac { -1 }{ 2 }  & \frac { 1 }{ 2 }  \\ 2 & -\frac { 1 }{ 2 }  & -\frac { 1 }{ 2 }  \\ -2 & 2 & 1 \end{matrix} \right] $$

## Validation in R

To validate the handsolution, I used the `matlib` package to get the inverse in r

```{r}
library(matlib)
a = matrix(c(1,3,1,
             0,2,1,
             2,2,1), nrow = 3, byrow = TRUE)
ai <- inv(a)
ai
```





