---
title: "Week_2 605 assignment"
author: "Salma Elshahawy"
date: "1/14/2020"
output: 
  pdf_document: 
    extra_dependencies: ["geometry"]
    latex_engine: xelatex
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem set_1

### (1) Show that "${ A }^{ T }A\quad \neq \quad A{ A }^{ T }$" in general.

```{r}
# creating a matrix A
A <- matrix(c(1,2,3,4,5,6,7,8,9), nrow = 3)
A
```
```{r}
# get A transpose and assign it to At
At <- t(A)
At
```

```{r}
# mutliply matrix A by it's transpose A*t(A)
left_side <- A %*% At
left_side
```

```{r}
# mutliply matrix transpose A by the original matrix t(A)*A
right_side <- At %*% A
right_side
```
```{r}
# check if the two sides are in equilibrium state
left_side == right_side
```

Both sides are not equal, so in general ${ A }^{ T }A\quad \neq \quad A{ A }^{ T }$



### (2) For a special type of square matrix A, we get ${ A }^{ T }A\quad =\quad A{ A }^{ T }$ . Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

_Please typeset your response using LaTeX/SWeave mode in RStudio. If you do it in paper, please either scan or take a picture of the work and submit it. Please en- sure that your image is legible and that your submissions are named using your first initial, last name, assignment and problem set within the assignment._

```{r}
# construct a diagonal matrix
# 
A <- matrix(c(1,0,0,0,5,0,0,0,9), nrow = 3)
A <- 5 * A # multiply by a scaler
A
```

```{r}
A%*%t(A) == t(A)%*%A
```

I Observed that when a matrix is symmetric, the matrix is equal to its transpose, ${ A }^{ T }A\quad =\quad A{ A }^{ T }$

The conditions:

  1- Be a square matrix $A_{ i\times j }\quad =\quad { A }_{ j\times i }^{ T }$
  
  2- The upper and lower triangle are equal to zero $\begin{matrix} { a }_{ ij } & 0 & 0 \\ 0 & { a }_{ ij } & 0 \\ 0 & 0 & { a }_{ ij } \end{matrix}$

## Problem set_2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

_Please submit your response in an R Markdown document using our class naming convention. You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems._

```{r}
a <- matrix(c(2,4,-2,
              1,-1,5,
              3,3,5), nrow = 3, byrow = TRUE)
a

get_UL <- function(a) {
  U = a
  L = diag(nrow(a))
  n = nrow(a)
  for (i in 1:n) {
    k = seq(2, n)
    for (j in k) {
      if(j > i) {
        # get the multiplier and add it to the L matrix
        s = U[[j,i]]/U[[i,i]]
        L[j,i] = s
        # reduce by reduction and shuffle to the U matrix
        U[j,] = U[j,] - s * U[i,]
      }
    }
  }
  return(list(U = U, L = L))
}

value <- get_UL(a)  
U_matrix <- value$U
U_matrix
L_matrix <- value$L
L_matrix
```







