---
title: "SElshahawy_Assign15"
author: "Salma Elshahawy"
date: "5/11/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question_1

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

(5.6,8.8),(6.3,12.4),(7,14.8),(7.7,18.2),(8.4,20.8)

```{r}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

reg <- lm(y~x)
summary(reg)
```

The regression line equation would be $y = 4.2471x - 14.8$


## Question_2

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form $( x, y, z )$. Separate multiple points with a comma.
$$f(x, y) = 24x - 6xy^2 - 8y^3$$
$f_x(x, y) = 24 - 6y^2$ and $f_y(x, y) = 24y^2 - 12xy$

To get $x$ and $y$ we solve the equations with zero

$24 - 6y^2 = 0$ then $y = \pm 2$ substitute in the 2nd equation to get the value of $x$
When $y = +2$ then $24y^2 - 12xy = 0$ then $x = -4$
When $y = -2$ then $24y^2 - 12xy = 0$ then $x = 4$

$f_(-4, 2) = 24\times (-4) - 6\times(-4)\times 2^2 - 8 \times 2^3 = -64$
$f_(4, -2) = 24\times 4 - 6\times4\times (-2)^2 - 8 \times (-2)^3 = 64$

Critical points $(-4, 2, -64)$ and $(4, -2, 64)$

Find maxima and minima 

$f_{xx} = 0$ , $f_{yy} = -12x-48y$, $f_{x,y} = -12y$

then $D(x,y) = f_{xx} f_{yy}-f^2_{xy} = -(-12y)^2 = -144y^2 < 0$ the any critical point is a saddle point.

## Question_3

A grocery store sells two brands of a product, the “house” brand and a “name” brand. The manager estimates that if she sells the “house” brand for x dollars and the “name” brand for y dollars, she will be able to sell $81−21x+17y$ units of the “house” brand and $40+11x−23y$ units of the “name” brand.

Step 1. Find the revenue function $R(x,y)$. 
Step 2. What is the revenue if she sells the “house” brand for \$2.30 and the “name” brand for $4.10?

$$R(x,y) = (81 - 21x + 17y)x + (40 + 11x - 23y)y$$
$$81x-21x^2+17xy+40y+11xy-23y^2$$

$$81x+40y+28xy-21x^2-23y^2$$
$$R(2.3, 4.1)=81\times 2.3+40\times 4.1+28\times 2.3\times 4.1-21\times (2.3)^2-23\times (4.1)^2 = 116.62$$


## Question_4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

let $x$: number of products in Los Angles, $y$: number of products produced in Denver.

Then $x + y = 96$ and $x = 96-y$

$$C(x,y) = C(96-y,y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$$
$$C_1(y) = \frac{1}{6} (96-y)^2 + \frac{1}{6} y^2 + 7\times (96-y) + 25y + 700$$
$$C_1(y) = \frac{1}{6}(y^2 - 192 y + 9216) + \frac{1}{6}y^2+672-7y+25y+700$$
$$C_1(y) = \frac{1}{6}y^2 - 32y+1536+\frac{1}{6}y^2+18y+1372$$
$$C_1(y) = \frac{1}{3}y^2 - 14y + 2908$$
$$C_1'(y) = \frac{2}{3}y-14$$

minimal value when $C_1'(y) = 0$ then $\frac{2}{3}y-14 = 0$ then $y = 21$ and $x = 96 - 21 = 75$

There should be 75 units produced in Los Angeles and 21 units produced in Denver.

## Question_5



















