---
title: "SElshahawy_Assign_13"
author: "Salma Elshahawy"
date: "4/28/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question_1

Use integration by substitution to solve the integral below.

$\int{4e^{-7x}dx}$

$$=\int { \frac { 4\times -7 }{ -7 }  } { e }^{ u }\quad du$$ 
$$=\frac { -4 }{ 7 } { e }^{ -7x }\quad +\quad c$$
where c is a constant.

## Question_2

Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt} = -\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function $N(t)$ to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

$$\frac{dN}{dt} = N'(t) = \frac{-3150}{t^4}-220$$
To find $N(t)$ we find the integration 

$$N(t) = \int{(\frac{-3150}{t^4}-220) dt}$$ 
$$N(t)= \frac{1050}{t^3}-220t+C$$
back substitute with the $N(t) = 6530$ where t = 1 to find C

$$6530= \frac{1050}{(1)^3}-220(1)+C$$
$$5700= C$$

The contamination level could be presented by

$$N(t)= \frac{1050}{t^3}-220t+5700$$

## Question_3

Find the total area of the red rectangles in the figure below, where the equation of the line is $f(x) = 2x - 9$

from the graph provided, we can see that each column have a width of 1 unit. The length can be found using the integration and substitute with the min and max value for the rectangle width. 

```{r}
eq = function(x) {
  return (2*x - 9)
}
area = integrate(eq, 4.5, 8.5)
print(area)
```

## Question_4

Find the area of the region bounded by the graphs of the given equations.

$$y_1 = x_1^2 - 2x_1-2$$
$$y_2 = x_2 + 2$$

```{r}
y_1 = function(x) {
  return(x**2 - 2*x - 2)
}

y_2 = function(x) {
  return(x + 2)
}

x_1 <- seq(-3, 6, 0.05)
plot(x_1, y_1(x_1), type='l', col="red", 
     xlab="", ylab="")
lines(x_1, y_2(x_1), col="green")
abline(h=0)
```

We need 4 point to find the are. Find roots of the quadratic function $f_1(x)$

```{r}
roots <- polyroot(c(-2, -2, 1))
```

the intersection of two functions should equal to zero

$$(x^2-2x-2)-(x+2)=0$$

$$x^2-3x-4=0$$


```{r}
intersects = polyroot(c(-4, -3, 1))
```

points are 

```{r}
points_x = c(intersects[1], roots, intersects[2])
points_x
```

```{r}
plot(x_1, y_1(x_1), type='l', col="red", 
     xlab="", ylab="")
lines(x_1, y_2(x_1), col="green")
points(points_x, y_1(points_x))
text(points_x, y_1(points_x), labels = c("a","b","c","d"), pos = 3)
abline(h=0)
```


```{r message=FALSE, warning=FALSE}
# Find individual parts
a1 <- integrate(y_2, lower=points_x[1], upper=points_x[4])
a2 <- integrate(y_1, lower=points_x[1], upper=points_x[2])
a3 <- integrate(y_1, lower=points_x[3], upper=points_x[4])
a4 <- integrate(y_1, lower=points_x[2], upper=points_x[3])

# Combine individual parts 
# Last integral is negative because it is below the x-axis, so it needs to be negated
area <- a1$value-a2$value-a3$value-a4$value
area
```



