---
title: "SElshahawy_Assign10"
author: "Salma Elshahawy"
date: "3/31/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if

### (a) 

he bets 1 dollar each time (timid strategy).

By applying the GAMBLER’S RUIN formula we can get the probability of winning the bet.

$$ P = \frac{(\frac{q}{p})^z - 1}{(\frac{q}{p})^M - 1} \\$$

```{r}
p = 0.4 # winning probability
q = 0.6 # losing probability
z = 1 # how many dollars he started with
M = 8 # when he stops
P = ((q/p)^z - 1) / ((q/p)^M - 1)  
P # probability to win the bet
```

### (b) 

he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

To win with bold strategy, the prisoner should put his wins dollars into the bet. So, if he bet with 1$ -> win will give him 2, 2 will give him 4, 4 will give 8 then he can stop and break free.To do that, he needs 3 successive wins in a row ( binomial distribution - two independent events like coin flip). We can do that using the `dbinom()` setting the probability of winning to 0.4

```{r}
dbinom(3, 3, 0.4)
```

## (c) 

Which strategy gives Smith the better chance of getting out of jail?

From calculations, it seems that the bold strategy gives  the prisoner a better chance  than the timid strategy to win this bet. 


