---
title: "SElshahawy_Assign10"
author: "Salma Elshahawy"
date: "3/31/2020"
output: 
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if
(a) he bets 1 dollar each time (timid strategy).

$$ P = \frac{1-(\frac{q}{p})^s}{1-(\frac{q}{p})^M} \\$$

```{r}
p = 0.4 # winning probability
q = 0.6 # losing probability
s = 1 # how many dollars he started
M = 8 # when he stops
P = (1-(q/p)^s) / (1-(q/p)^M)  
P # probability to win the bet
```


(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).


(c) Which strategy gives Smith the better chance of getting out of jail?

