---
title: "Final_Exam part#1"
author: "Salma Elshahawy"
date: "2020-05-15"
output: 
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    number_section: false
    toc_float:
      collapsed: true
      smooth_scroll: true
    df_print: paged
  pdf_document:
    extra_dependencies: geometry
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Your final is due by the end of the last week of class. You should post your solutions to your GitHub account or RPubs. You are also expected to make a short presentation via YouTube  and post that recording to the board. This project will show off your ability to understand the elements of the class.

## Problem 1

Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.  

```{r}
set.seed(123)
N <- runif(1, 8, 1000)
n <- 10000
X <- runif(n, min = 0, max = N)

```

Then generate a random variable Y that has 10,000 random normal numbers with a mean of $μ=σ=(N+1)/2$.

```{r}
set.seed(123)
me <- (N+1)/2
sd <- me
Y <- rnorm(n, mean = me, sd = sd)

```


## Probability {.tabset}

Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.  Interpret the meaning of all probabilities.

```{r}
(x <- median(X))
(y <- quantile(Y, 0.25)[[1]])
```

### P(X>x | X>y)

$$P(X>x | X>y) = \frac{P(X>145.0453) \quad  P(X>48.85452) }{P(X>48.85452)}$$


```{r}
(p_a <- sum(X>x & X>y)/n)
(p_xy <- sum(X>y)/n) 
(p_a/p_xy)
```
0.6 is the probability that X is **greater than** its median given that X is **greater than** the first quartile of Y.

### P(X>x , Y>y)

$$P(X>x , X>y) = {P(X>145.0453) . P(Y>48.85452) }$$

```{r}
(p_b <- sum(X>x & Y>y)/n)
```

0.36 is the probability that X and Y are **greater than** all possible x and y. 

### P(X<x | X>y)

$$P(X<x | X>y) = \frac{P(X<145.0453) . P(X>48.85452) }{P(X>48.85452)}$$


```{r}
(p_c <- sum(X<x & X>y)/n)
(p_xy <- sum(X>y)/n) 
(p_c/p_xy)
```

0.39 is the probability of X **less than** its median **and greater than** the first quantile of Y.


### Investigating the formula 

Investigate whether $P(X>x \quad and \quad  Y>y)=P(X>x)P(Y>y)$ by building a table and evaluating the marginal and joint probabilities.

```{r}
(res <- matrix(c(sum(X>x & Y<y),sum(X>x & Y>y), sum(X<x & Y<y),sum(X<x & Y>y)),  ncol = 2, nrow = 2))
res <- cbind(res,c(res[1,1] + res[1,2], res[2,1] + res[2,2]))
res <- rbind(res,c(res[1,1] + res[2,1], res[1,2] + res[2,2], res[1,3] + res[2,3]))
(results <- as.data.frame(res))
results
colnames(results) <- c("X>x", "X<x", "total")
rownames(results)  <- c("Y<y", "Y>y", "total")

results
```

## Probability matrix {.tabset}

```{r}
(prob_tab <- results/n)
```

Check $P(X>x \quad and \quad  Y>y)=P(X>x)P(Y>y)$
Check the right side : $P(X>x)P(Y>y)$ from the table we get

```{r}
(round(0.5*0.75, 2))
```

Check the leftside: $P(X>x \quad and \quad  Y>y)$ from the table = 0.369 ~ 0.38

Since the results are so similar we can conclude that both X and Y are independent variable

Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test.  What is the difference between the two? Which is most appropriate?

### Fisher test

```{r}
fisher.test(results,simulate.p.value=TRUE)
```

### Chai-Test

```{r}
chisq.test(results, correct=TRUE)
```

### Comparasion

**"Fisher's exact test"** is a way to test the association between two categorical variables when you have small cell sizes (expected values less than 5). While, **Chi-square** test is used when the cell sizes are expected to be large. If the sample size is small (or expected cell sizes <5),Fisher's exact test should be used. Otherwise, the two tests will give relatively the same answers. With large cell sizes, their answer should be very similar.

## Modeling, Kaggle competition

  - [Github repo - exam](https://github.com/salma71/MSDS_SP2020/tree/master/Spring2020/computational_mod_605/Final_exam)
  - [Kaggle account - username: salmaeng](https://www.kaggle.com/salmaeng)
  - [House prices - kaggle advance regression model - part#2](https://rpubs.com/salmaeng/kaggle_houseprices)
  - [Presentation]()